---
title: "SD_dataset"
author: |
    | Katarzyna Otko s202872
    | Alvils Sture s202586
date: "October 2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

## 1. A description of the dataset

### 1.1. What is our data about?

The dataset of our choice is a result of a **Speed Dating Experiment** that was conducted in years 2002 - 2004 by professors of Colombia University for their paper entitled *Gender Differences in Mate Selection: Evidence From a Speed Dating Experiment*. The goal of this two-year experiment was to discover what features and factors play the most important role during first minutes of a date. In others words, what makes us want to see the other person again?

Source of the dataset: https://data.world/annavmontoya/speed-dating-experiment

Overall, there were 21 speed dating events (waves), with participants ranging from 12 to 44 in each, which in overall sums up to 551 people. Those subjects were drawn from students in graduate and professional schools at Columbia University. Each date lasted 4 minutes.

The dataset also includes questionnaire data gathered from participants at different stages of the process. Attendees answered questions about their dating habits, self-perception, beliefs on what others may find attractive in a mate, lifestyle, demographics and many more. The process consisted of 4 steps:

* Signup (Time 1) - filling out a survey in order to register for the event;
* Answering questions half way through meeting all potential dates during the night of the event;
* 1st Followup (Time 2) - filling out a survey the day after participating in the event;
* 2nd Followup (Time 3) - filling out a survey 3-4 weeks after participants had been sent their matches.

### 1.2. Summary of previous analysis of the dataset

We would like to briefly summarize the previously mentioned paper *Gender Differences in Mate Selection: Evidence From a Speed Dating Experiment* written by Raymond Fisman, Sheena S. Iyengar, Emir Kamenica and Itamar Simonson.




```{r, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)
library(tidyr)
setwd('C:/Users/katin/Desktop/Folder/STUDIA/DTU/Semestr I/Intro to ML/Project I')
setwd("/home/nomow/Documents/DTU/Intro_to_machine_learning/ML_speed_dating")
SD <- read.csv('Speed Dating Data.csv')

# numdim(SDber of rows and columns
dim(SD)

# number of women
length(unique(SD$iid[which(SD$gender == 0)])) # 274

# number of men
length(unique(SD$iid[which(SD$gender == 1)])) # 277

274 + 277
```


## Missing values
```{r}
NAs <- sapply(SD, function(x) sum(is.na(x)))
sort(NAs[which(NAs > 0)]) 

#filling one missing value in last id row
SD[which(is.na(SD$id)), 1:2] <- 22

# filling 10 missing values in pid columns
SD[which(is.na(SD$pid)), 1:15] # partner's id - 7
SD[which(SD$id == 7 & SD$wave == 5), 1:2] # we have to fill these 10 NAs with 128
SD[which(is.na(SD$pid)), 'pid'] <- 128


# filling missing values in columns <attribute_name>1_1
DF <- SD[, 70:75]
df <- DF[!complete.cases(DF),]
dim(df); sum(rowSums(df, na.rm = T) == 100) # so there are 42 rows where we can impute 0s
df[which(df[,1] + df[,2] + df[,3] == 100),]
df[which(df[,1] + df[,2] + df[,3] == 100), c(4:6)] <- 0

df[which(df[,1] + df[,2] + df[,3] + df[,4] == 100),]
df[which(df[,1] + df[,2] + df[,3] + df[,4] == 100), c(5:6)] <- 0

df[which(df[,1] + df[,2] + df[,3] + df[,4] +df[, 5] == 100),]
df[which(df[,1] + df[,2] + df[,3] + df[,4] +df[, 5] == 100), 6] <- 0

DF[!complete.cases(DF),] <- df
SD[,70:75] <- DF

df <- subset(SD, !duplicated(SD$iid))
df[,1]
67/nrow(df)
NAs <- sapply(df, function(x) sum(is.na(x))); NAs
NAs_sum <- sum(NAs); NAs_sum

# adding one column with explanation for race column (matching index with race names)
race_idx <- unique(SD$race)
race_val <- c('Asian', 'European', 'Other', 'Latino', 'Black', NA)
SD$race_explained <- race_val[match(SD$race, race_idx)]

# adding one column with explanation for field_cd column (matching index with race names)
# DISCUSS WITH ALVILS IMPUTING DATA INTO field_cd as 9 (because field is Operations Research)
field_idx <- c(1:18, NA)
field_val <- c('Law', 'Math', 'Social Science, Psychologist', 'Medical Science/Pharmaceuticals/Bio Tech',
               'Engineering', 'English/Creative Writing/ Journalism', 'History/Religion/Philosophy',
              'Business/Econ/Finance', 'Education, Academia', 'Biological Sciences/Chemistry/Physics', 
              'Social Work', 'Undergrad/undecided', 'Political Science/International Affairs',
              'Film', 'Fine Arts/Arts Administration', 'Languages', 'Architecture', 'Other', 'Other')
SD$field_explained <- field_val[match(SD$field_cd, field_idx)]

# converting income from string to numeric
SD$income <- as.numeric(gsub(',', "", SD$income, fixed = T))
  
sum(is.na(SD$income))

unique(SD$field_cd)

summary(SD[SD$wave >= 6 & SD$wave <= 9,129:134])
# Waves 6 - 9:
# attr4_1 - shar4_1 have values between 0 and 10
# attr2_1 - shar2_1 OK
# attr1_2 - shar1_2 OK

```

```{r}
# Importance of each attribute by gender
attr <- subset(SD, !duplicated(SD$iid), select = c(3, 70:75)) %>%
  filter(!is.na(attr1_1))  %>% # 7 people with missing values
  group_by(gender) %>%
  summarize(Attractiveness = mean(attr1_1),
            Sincerity = mean(sinc1_1),
            Intelligence = mean(intel1_1),
            Fun = mean(fun1_1, na.rm = T),
            Ambition = mean(amb1_1, na.rm = T),
            'Shared interests' = mean(shar1_1, na.rm = T))

data_long <- pivot_longer(attr, !gender) # manipulating data to  plot

data_long %>% ggplot(aes(x = name, y = value, fill = factor(gender))) + 
  geom_bar(stat = 'identity', position = 'dodge') +
  xlab('Attribute') + ylab('Importance') + ggtitle('Importance of each attribute by gender') +
  geom_text(aes(label = round(value, 2)), position=position_dodge(width=0.9), vjust = -0.5, size = 3.2) +
  scale_fill_discrete(name="Gender",
                         breaks=c(0, 1),
                         labels=c("Women", "Men")) + # Editing legend
  theme_minimal() + # Changing theme
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0), size = 12),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0), size = 12)) # Centers title
```

```{r}
# Age analysis
sum(is.na(SD$age))
SD[is.na(SD$age), 1:10]

age_df <- subset(SD, !duplicated(SD[,1])) %>%
  filter(!is.na(age)) %>%
  group_by(wave, gender) %>%
  summarize(Average_age = mean(age))

SD %>% nrow()
nrow(SD)

age_df$gender <- ifelse(age_df$gender == 0, 'Women', 'Men')

# Mean age per wave
age_df %>% ggplot(aes(x = wave, y = Average_age, fill = gender)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_discrete(name = "Gender") 

age_df <- subset(SD, !duplicated(SD$iid), select = c(iid, gender, age)) %>%
  filter(!is.na(age)) %>%
  mutate(mean = mean(age))
age_df$gender <- ifelse(age_df$gender == 0, 'Women', 'Men')

# Histogram of age
max(unique(age_df$age)) - min(unique(age_df$age)) # number of bins
age_df %>% ggplot(aes(x = age)) + 
  geom_histogram(bins = 37, fill = 'lightgrey', position = 'identity', alpha = .7) +
  geom_vline(aes(xintercept = mean), col = 'red', linetype = 'dashed')
```

```{r}
# Field analysis
field_df <- subset(SD, !duplicated(SD$iid)) %>%
  filter(!is.na(field_cd)) %>%
  group_by(field_explained, gender) %>%
  summarize(field_sum = n())

field_df$gender <- ifelse(field_df$gender == 0, 'Women', 'Men')

field_df %>% ggplot(aes(x = field_explained, y = field_sum, fill = gender)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  coord_flip()

```

```{r}
# Income
income_df <- subset(SD, !duplicated(SD$iid)) %>%
  filter(!is.na(income))

income_df$gender <- ifelse(income_df$gender == 0, 'Women', 'Men')

income_df %>% ggplot(aes(x = income/1000, fill = gender)) + 
  geom_histogram(bins = 15)

```

```{r}
# Purpose
goal_df <- subset(SD, !duplicated(SD$iid)) %>%
  filter(!is.na(goal)) %>%
  group_by(goal, gender) %>%
  summarise(count = n())

goal_df$gender <- ifelse(goal_df$gender == 0, 'Women', 'Men')

goal_idx <- unique(goal_df$goal)
goal_val <- c('Seemed like a fun night out', 'To meet new people', 'To get a date', 
              'Looking for a serious relationship', 'To say I did it',	'Other')
goal_df$goal_explained <- goal_val[match(goal_df$goal, goal_idx)]

goal_df %>% ggplot(aes(x = goal_explained, y = count, fill = gender)) + 
  geom_bar(stat = 'identity', position = 'dodge') +
  coord_flip()

```

```{r}
# Importance of features for men/women

```

## PCA

```{r}
colnames(SD)[51:67]
Y <- subset(SD, !duplicated(SD$iid), select = c(51, 56:60, 63, 65, 66))
Y <- t(apply(Y, 1, '-', colMeans(Y, na.rm = T)))

s <- svd(Y[complete.cases(Y),])
diagS <- s$d
rho <- diagS^2/sum(diagS^2)

threshold = 0.9


xlimits <- c(1, ncol(Y)); 
plot(rho,
     type='o',
     main="Variance explained by principal components",
     xlab="Principal components", 
     ylab="Variance explained",
     xlim=xlimits,
     ylim=c(0,1),
     col='blue')
lines(cumsum(rho), type='o', col='orange')
lines(xlimits, c(threshold, threshold), lty='dashed')

legend("right", # Define position
       legend=c("Cumulative", "Individual", "Threshold"), # Set strings for legend
       col=c("orange", "blue", "black"), lty=c(1,1,2), # Match appereance of lines
       cex=1.5, bg='lightblue') # Setup how the box looks (cex controls size)
```
## Scree plot

eigenv = diagS^2
xlimits <- c(1, ncol(Y)); 
plot(eigenv,
     type='o',
     main="Scree plot",
     xlab="Principal component", 
     ylab="Eigenvalue",
     xlim=xlimits,
     col='blue')
     
## removes all nan values
Y <- subset(SD, !duplicated(SD$iid), select = c(51, 56:60, 63, 65, 66))
match = subset(SD, !duplicated(SD$iid), select = c(13))
match = match[complete.cases(Y),]

Y = Y[complete.cases(Y),]

## cos dist
match_Y = Y[which(match == 1), ]
centr_match_Uy = matrix((colMeans(match_Y) / sqrt(sum(colMeans(match_Y)^2))))

no_match_Y = Y[which(match == 0), ]
centr_no_match_Uy = matrix((colMeans(no_match_Y) / sqrt(sum(colMeans(no_match_Y)^2))))

uY = data.matrix(Y/sqrt(rowSums(Y^2, na.rm=T)))

##
score_no_match = (uY) %*% centr_no_match_Uy
score_match = (uY) %*% centr_match_Uy


